\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-09-30 gwnnDraft02.tex
Modified conclusion and added scenario details.

\begin{equation}\label{eqn:Beta}

@@

The ratio of $\beta$ and the Hubble constant $H$, at the time of nucleation, controls both the strength and peak frequency of the GW signal. Using Eq.~(\ref{eqn:Beta}) it can be expressed explicitly as, 

@@

Due to the lack of a general analytic QCD potential, it is not possible to use Eq.~(\ref{eqn:BH}) to calculate $\beta/H$. There are dimensional arguments~\cite{Hogan:1984hx,Hogan:1986qda} which predict $\beta/H \sim 4 \textrm{Log}(M_{p}/ \Lambda_{QCD,i})$, although these arguments are based on assumptions concerning the potential. In more recent work, some authors~\citep{Helmboldt:2019pan,Bai:2018dxf} have attempted to estimate it using first-order chiral effective theories and the Polyakov-Nambu-Jona-Lasinio (PNJL) models. These studies claim a large range of values and thus a consensus has not yet been reached.  Under these circumstances, we take the optimistic case in which $\beta/H$ is of order $\mathcal{O}(10)$.

@@

Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in figure~(\ref{fig::Nnatural}). Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, more detail in the construction of these curves is given in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sectors contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density. This is not the case if we consider the more generalized reheating scenarios of section~\ref{sec:constraints}. 
\draftnote{I think we need more details on the scenarios that go into making those plots.} Assuming all else equal, we use the temperature schemes from Eq.~(\ref{eqn:energyDensityAllowed}) to produce the plots in figure~(\ref{fig:Haa}). In all cases, the summed GW signal is detectable by one or more proposed interferometers. When relaxing the assumptions on $\beta/H$, the scenarios in figure~(\ref{fig:Haa}) are still detectable for values ranging between $\mathcal{O}(1)$ and $\mathcal{O}(100)$. The frequency dependence in Eq.~(\ref{eqn::spectralshape}) takes the form of $f/f_{p}$, this causes a cancellation between the redshifting factors. As multiple sectors phase transition at different times, and therefore temperatures, the peaks will shift relative to each other, purely from the linear temperature dependence of the peak frequency $f_{p} \sim T_{\gamma}$.  This is seen in figure~(\ref{fig:Haa}), where the spectrum peaks are shifted causing a peak broadening of the convoluted spectrum. The broadening can be substantial if the hidden sectors transition between a large gap of time (temperature). Eventually, a temperature limit will be reached at which two (or multiple) peaks, if the amplitudes are comparable. 

@@

Here, we examined several scenarios (including $N$naturalness) that involve multiple hidden sectors and calculated the GW profiles present. Our GW projections unambiguously demonstrate that scenarios with multiple hidden sector SFOPTs can begin to be probed by next generation experiments. Further, our results provide a framework for understanding and using GW signals in two opposite ways: first as a unique signal for specific theories featuring multiple SFOPTs and also as a way to explain signals that deviate from a standard power law spectrum.

@@

Shifting to the latter scenario, should a stochastic GW background be detected, deviations from a standard power law curve would indicate a more complicated hidden sector.  Explicitly, a broadening or distortion of the signal around the peak frequency could point to a multi-SFOPT scenario --- and gently guide us in the direction of multiple hidden sectors.

>>>>

@@

The ratio of $\beta$ and the Hubble constant, at the time of nucleation, $H$ controls the strength of the GW signal,

@@

Due to the lack of a general analytic QCD potential, it is not possible to use Eq.~(\ref{eqn:BH}) to calculate $\beta/H$. There are dimensional arguments~\cite{Hogan:1984hx,Hogan:1986qda} which predict $\beta/H \sim 4 \textrm{Log}(M_{p}/ \Lambda_{QCD,i})$, although these arguments are riddled with assumptions on the potential. In more recent work, some authors~\citep{Helmboldt:2019pan,Bai:2018dxf} have attempted to estimate it using first-order chiral effective theories and the Polyakov-Nambu-Jona-Lasinio (PNJL) models. These studies claim a large range of values and thus a consensus has not yet been reached.  Under these circumstances, we take the optimistic case in which $\beta/H$ is of order $\mathcal{O}(10)$.

@@

Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in figure~(\ref{fig::Nnatural}). Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, more detail in the construction of these curves is given in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sectors contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

This is not the case if we consider the more generalized reheating scenarios of section~\ref{sec:constraints}. Once the restriction that small vevs are preferentially reheated has been lifted, we can explore a landscape of standard sectors with much heavier particles (specifically quarks) and, as shown in section~\ref{sec:dQCD}, the SFOPT these entail. The key phenomenological constraint on such sectors is $\Delta N_{eff}$, giving us a maximum allowed temperature ratio (when compared to the SM) for each reheated hidden sector: Eq.~(\ref{eqn:energyDensityAllowed}) shows the maximum temperature ratios for specific numbers of additional hidden sectors. There are four scenarios that we examine here:
\item \textbf{Maximized signal:} A single additional heavy hidden sector which saturates all current experimental bounds. In the $N$naturalness framework this is equivalent to reheating a standard sector with $i\sim 10^{16}$ up to the maximum allowed temperature ratio.
\item \textbf{Large split scenario:} A scenario where two additional hidden sectors have been reheated --- these sectors have higgs vevs that are split by a factor of $\sqrt{10^3}$, leading to a large difference in the frequency of their gravitational wave signals.  In the $N$naturalness framework this is equivalent to reheating two standard sectors, one with $i\sim 10^{12}$ and another with $i\sim 10^{15}$ up to the maximum allowed temperature ratio.
\item \textbf{Medium split scenario:} Similar to the previous case: these sectors have higgs vevs that are split by a factor of $\sqrt{10}$, resulting in a much smaller difference in the frequency of their gravitational wave signals.  In the $N$naturalness framework this is equivalent to reheating two standard sectors, one with $i\sim 10^{12}$ and another with $i\sim 10^{13}$ up to the maximum allowed temperature ratio.
\item \textbf{Five sector scenario:} Five sectors are reheated to the maximum allowed temperature ratio, each with vevs that are $\sim \sqrt{3}$ larger than the previous sector.
The GW results of these cases are presented in figure~(\ref{fig:Haa}). In all cases, the summed GW signal is detectable by one or more proposed interferometers. When relaxing the assumptions on $\beta/H$, the scenarios in figure~(\ref{fig:Haa}) are still detectable for values ranging between $\mathcal{O}(1)$ and $\mathcal{O}(100)$ The frequency dependence in Eq.~(\ref{eqn::spectralshape}) takes the form of $f/f_{p}$, this causes a cancellation between the redshifting factors. As multiple sectors phase transition at different times, and therefore temperatures, the peaks will shift relative to each other, purely from the linear temperature dependence of the peak frequency $f_{p} \sim T_{\gamma}$.  This is seen in figure~(\ref{fig:Haa}), where the spectrum peaks are shifted causing a peak broadening of the convoluted spectrum. The broadening can be substantial if the hidden sectors transition between a large gap of time (temperature). Eventually, a temperature limit will be reached at which two (or multiple) peaks, if the amplitudes are comparable. 

@@

Here, we examined several scenarios (including $N$naturalness) that involve multiple hidden sectors and calculated the GW profiles present. Our GW projections demonstrate that although vanilla $N$naturalness is not projected to be detectable in the near future, more generalized scenarios with multiple hidden sector SFOPTs are in an observable region and will begin to be probed by next generation space experiments. Both cases feature important parts of their GW signals in the void between frequencies detectable by pulsar timing arrays and space-based interferometers --- providing theoretical impetus for new experiments capable of probing this region of frequency space.

Further, our results provide a framework for understanding and using GW signals in two opposite ways: first as a unique signal for specific theories featuring multiple SFOPTs and also as a way to explain signals that deviate from a standard power law spectrum.

@@

Shifting to the latter scenario, should a stochastic GW background be detected, deviations from a standard power law curve would indicate a more complicated hidden sector. Individual SFOPT are understood to create GW that are assumed to follow an approximate power law. The multiple transitions that occur in the models outlined here create signals that do not respect this trend: although the individual GW do obey approximate power laws, their sum does not --- leading to a unique signal indicating so-called dark complexity. Explicitly, a broadening or distortion of the signal around the peak frequency could point to a multi-SFOPT scenario and gently guide us in the direction of multiple hidden sectors.
